Structured Versus Decorated Cospans

### 2020-12-17 cospans.tex
changed date on top of file
changes in big diagrams on pp. 13-14, and text after second one.

% Structured versus decorated cospans
% John Baez, Kenny Courser and Christina Vasilakopoulou
% in LaTeX
% 2020/12/16

@@

The associator for horizontal composition
is formed as follows: for three composable horizontal 1-cells $M_1=(a\to m_1\leftarrow b,x_1\in F(m_1))$, $M_2=(b\to m_2\leftarrow c,x_2\in F(m_2))$ and $M_3=(c\to m_3\leftarrow d,x_3\in F(m_3)),$
the isomorphism $(M_3\odot M_2)\odot M_1\cong M_3\odot (M_2\odot M_1)$ is the isomorphism of cospans 
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0.5) {$a$};
\node (A') at (0,-0.5) {$a$};
\node (B) at (1.5,0.5) {$m_1+_{b}(m_2 +_{c} m_3)$};
\node (C) at (3,0.5) {$d$};
\node (C') at (3,-0.5) {$d$};
\node (D) at (1.5,-0.5) {$(m_1+_{b} m_2)+_{c} m_3$};
\node (E) at (5.5,0.5) {$(x_3 \odot x_2) \odot x_1 \in F(m_1+_{b} (m_2 +_{c} m_3))$};
\node (F) at (5.5,-0.5) {$x_3 \odot (x_2 \odot x_1) \in F((m_1+_{b} m_2)+_{c} m_3)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$$} (B)
(C) edge node[above]{$$} (B)
(A) edge node[left]{$1_a$} (A')
(C) edge node[right]{$1_d$} (C')
(A') edge node {$$} (D)
(C') edge node {$$} (D)
(B) edge node [left] {$\stackrel{\kappa}{\cong}$} (D);
\end{tikzpicture}
\]
together with an isomorphism $F(\kappa)((x_3 \odot x_2) \odot x_1) \cong (x_3 \odot (x_2 \odot x_1))$ between the decorations 
\begin{equation}\label{eq:decoiso}
\scalebox{.7}
{\begin{tikzcd}[ampersand replacement=\&]
\& F(m_1){\times}(F(m_2){\times}F(m_3))\ar[dddrr,phantom,"\stackrel{\cref{eq:omega}}{\cong}"]\ar[r,"1\times\phi_{m_2,m_3}"]\ar[ddd,"\alpha"description] \& F(m_1){\times}F(m_2{+}m_3)\ar[r,"1\times F(\psi)"]\ar[dr,"\phi_{m_1,m_2+m_3}"'] \& F(m_1){\times}F(m_2{+_c}m_3)\ar[d,phantom,"\stackrel{\cref{eq:pseudonaturality}}{\cong}"]\ar[r,"\phi_{m_1,m_2+_cm_3}"] \& F(m_1{+}(m_2{+_c}m_3))\ar[r,"F(\psi)"]\ar[dddr,phantom,"\cong"] \& F(m_1{+_b}(m_2{+_c}m_3))\ar[ddd,"F(\kappa)"] \\
\one\ar[ur,"x_1\times(x_2\times x_3)"]\ar[d,equal] \&\&\& F(m_1{+}(m_2{+}m_3))\ar[ur,"F(1+\psi)"']\ar[d,"F\alpha"] \& \\
\one\ar[dr,"(x_1\times x_2)\times x_3"']\&\&\& F((m_1{+}m_2){+}m_3)\ar[dr,"F(\psi+1)"]\ar[d,phantom,"\stackrel{\cref{eq:pseudonaturality}}{\cong}"] \& \\
\& (F(m_1){\times}F(m_2)){\times}F(m_3)\ar[r,"\phi_{m_1,m_2}\times1"'] \& F(m_1{+}m_2){\times}m_3\ar[ur,"\phi_{m_1+m_2,m_3}"]\ar[r,"F(\psi)\times1"'] \& F(m_1{+_b}m_2){\times} F(m_3)\ar[r,"\phi_{m_1{+_b}m_2,m_3}"'] \& F((m_1{+_b}m_2){+}m_3)\ar[r,"F\psi"'] \& F((m_1{+_b}m_2){+_c}m_3)
\end{tikzcd}}
\end{equation}
where $a$ are the associators of $\Cat$ and $\A$ and $\kappa$ is the canonical isomorphism between two colimits of the same diagram in $\A$. Similarly the right and left unitors can be constructed, and all data satisfy the axioms of a double category---essentially breaking down to those of the cospan double category of $\A$ and axioms of the lax monoidal pseudofunctor $F$. For full details of these arguments, see \cite[Theorem~4.1.1]{CourserThesis}.
\end{proof}

@@

\&\& F(m_1+n_1+m_2+n_2)\ar[d,"F\psi"] \\

@@

\one \ar[ddrr,dashed,"{(x_2\odot x_1) \otimes (y_2\odot y_1)}"']\ar[r,"x_1{\times}x_2{\times}y_1{\times}y_2"] \& F(m_1){\times}F(m_2){\times}F(n_1){\times}F(n_2)\ar[r,"\phi_{m_1,m_2}\times\phi_{n_1,n_2}"] \& F(m_1{+}m_2){\times}F(n_1{+}n_2) \ar[d, "F\psi\times F\psi"] \\

@@

\& F(m_1)\times F(n_1)\times F(m_2)\times F(n_2)\ar[dd,"1\times\beta\times1"']\ar[ddrr,phantom,"\stackrel{\cref{eq:thetaa}}{\cong}"]\ar[rr,"\phi_{m_1,n_1}\times\phi_{m_2,n_2}"] \&\& F(m_1+n_1)\times F(m_2+n_2)\ar[rr,"\phi_{m_1+n_1,m_2+n_2}"] \&\& F(m_1+n_1+m_2+n_2)\ar[dd,"F\psi"]\ar[ddll,"F(1+\beta+1)"'] \\

@@

where the left-hand side isomorphism combines pseudoassociativity \cref{eq:omega} and braided monoidal structure \cref{eq:braidedpseudofun} of the pseudofunctor $F$ as seen in detail in the appendix, and the right hand side follows from the universal property of colimits and pseudofunctoriality of $F$.
Similarly, for units $U_{a+b}$ and $U_a\ot U_b$ it is easy to see that as cospans they are both the identity $1_{a+b}:a+b\to a+b\leftarrow a+b:1_{a+b}$, and there is a canonical isomorphism between their trivial decorations \cref{eq:UFCsp}, namely $F(!_{a+b})\phi_0\cong\phi_{a,b}(F(!_a)\times F(!_b))(\phi_0\times\phi_0)$ via the structure isomorphisms for $(\phi,\phi_0)$.

@@

Building towards an isomorphism of the above described double categories, we now define a double functor
\begin{displaymath}
\lE=(\lE_0,\lE_1)\maps _L\lCsp(\inta F) \longrightarrow F\lCsp
\end{displaymath}
which on vertical categories acts as the identity $\lE_0 = \id_{\A}$. Given a horizontal 1-cell in $_L\lCsp(\inta F)$ as in \cref{eq:scsphor1cell}, $\lE_1$ maps it to the decorated cospan $i\maps a\to m\leftarrow b:o$ with decoration $x\in F(m)$,
whereas a 2-morphism of $L$-structured cospans as in \cref{eq:2morphismLCspintaF} is mapped to the underlying map of cospans in $\A$
\begin{displaymath}
\begin{tikzcd}
a\ar[r,"i"]\ar[d,"f"'] & m\ar[d,"k"] & b\ar[l,"o"']\ar[d,"g"] \\
a'\ar[r,"i'"']& m' & b'\ar[l,"o'"]
\end{tikzcd}
\end{displaymath}
in $\A$ along with the map $\tau\maps Fk(x)\to x'$ in $F(m')$ coming from the middle arrow above, namely an $F$-decorated cospan 2-morphism. 
With these mappings, $\lE$ is indeed a double functor as in \cref{def:doublefun}. First of all, both $\lE_0$ and $\lE_1$ are functors that appropriately commute with sources and targets. Moreover, there exist globular 2-isomorphisms $\lE(N) \odot \lE(M) \simrightarrow\lE(N \odot M)$, $U_{\lE m} \simrightarrow\lE(U_m)$
for any pair of composable horizontal 1-cells $M=\cspn{(a,\bot_a)}{(m,x)}{(b,\bot_b)}{}{}$ and $N=\cspn{(b,\bot_b)}{(n,y)}{(c,\bot_c)}{}{}$ and any object $m$ of $_L\lCsp(\inta F)$. Explicitly, $\mathbb{E}(N)\odot \mathbb{E}(M)$ is given as in \cref{eq:dcospanscomposition} via the pushout cospan and decoration
\begin{equation}\label{eq:1}
 \begin{tikzcd}
  & m+_bn & \\
  a\ar[ur] && b,\ar[ul]
 \end{tikzcd}\qquad
  \one\xrightarrow{x\times y}F(m)\times F(n)\xrightarrow{\phi_{m,n}}F(m+n)\xrightarrow{F(\psi)}F(m+_bn)
\end{equation}
whereas first composing $M$ and $N$ in $_L\lCsp(\inta F)$ using pushouts in $\inta F$ as in \cref{eq:globalpushout} produces 
\begin{displaymath}
 \begin{tikzcd}[column sep={.7in,between origins}]
 && (m+_b n,F(v_m)x+_{(\bot_{m+_bn})}F(v_n)y) && \\
 & (m,x) \ar[ur] && (n,y)\ar[ul] \\
 (a,\bot_a)\ar[ur] && (b,\bot_b)\ar[ur]\ar[ul] && (c,\bot_c)\ar[ul]
 \end{tikzcd}
\end{displaymath}
where $v_m,v_n$ are the canonical maps into the pushout.
Under $\lE$ this is mapped to the underlying $\A$-cospan $\cspn{a}{m+_bn}{b}{}{}$ with decoration $F(v_m)x+F(v_n)y$ in the fiber $F(m+_bn)$, since a pushout over an initial object is really a coproduct. That decoration is isomorphic to that of \cref{eq:1} due to the following diagram 
\begin{equation}\label{eq:isomorphismdiag}
 \begin{tikzcd}
F(m)\times F(n)\ar[rr,"\phi_{m,n}"]\ar[d,"F(v_m)\times F(v_n)"']\ar[dr,phantom,"\cong"] && F(m+n)\ar[d,"F(\psi)"]\ar[d,bend right=50, phantom,"\cong"']\ar[dl,"F(v_m+v_n)"description] \\
F(m+_bn)\times F(m+_bn)\ar[r,"\phi_{m+_bn,m+_bn}"']\ar[rr,dashed,bend right=15] & F\left((m+_bn)+(m+_bn)\right)\ar[r,"F(\nabla)"'] & F(m+_bn)
 \end{tikzcd}
\end{equation}
where the dashed bottom composite is the fiberwise coproduct in $F(m+_bn)$ as in \cref{eq:explicitstructure1}. The left-hand side isomorphism is \cref{eq:pseudonaturality} from pseudonaturality of $\phi$ whereas the right-hand side isomorphism follows from universal properties and pseudofunctoriality of $F$.
Regarding identities, the identity horizontal 1-cell $U_m$ in $_L\lCsp(\inta F)$ is $\cspn{(m,\bot_m)}{(m,\bot_m)}{(m,\bot_m)}{}{}$ with $1_m$ as the $A$-component and isomorphisms between initial objects in the fibers. Thus $\lE(U_m)$ is the identity cospan in $A$ together with the `initial decoration' $\bot_m\in F(m)$, and so is $U_{\mathbb{E}(m)}$ as in \cref{eq:UFCsp}.

@@

Since both double categories are symmetric monoidal under the running assumptions (\cref{DC,thm:SC}), we now show that this double isomorphism is a symmetric monoidal one, namely that $\lE=(\lE_0,\lE_1)$ is a symmetric strong monoidal double functor as per \cref{defn:monoidal_double_functor}. Indeed, $\lE_0= \id_\A$ is a symmetric monoidal functor trivially. For $\lE_1$, given two structured cospans $M=(a,\bot_a)\to(m,x)\leftarrow(b,\bot_b)$ and $M'=(a',\bot_{a'})\to(m',x')\leftarrow(b',\bot_{b'})$ in $_L\lCsp(\inta F)$, we compute the appropriate decorated cospans
\begin{align*}
\lE(M)\ot\lE(M')&\;\stackrel{\cref{eq:tensordecoration}}{=}\;a+a'\xrightarrow{i+i'}m+m'\xleftarrow{o+o'}b+b' \textrm{ with }\phi_{m,m'}(x,x')\in F(m+m')
%\one\xrightarrow{x\times x'}F(m)\times F(m')\xrightarrow{\phi_{m,m'}}F(m+m')
\\
\lE(M\ot M')&\;\stackrel{\cref{eq:tensorstuctured}}{=}\;\lE\Big((a+a',\bot_{a+a'})\to\underbrace{(m+m',F(\iota_m)x+F(\iota_{m'})x')}_{\cref{eq:fibrewisecoprod}}\leftarrow(b+b',\bot_{b+b'})\Big) \\
&=a+a'\xrightarrow{i+i'}m+m'\xleftarrow{o+o'}b+b' \textrm{ with }F(\iota_m)x+F(\iota_{m'})x'\in F(m+m')
\end{align*}
The cospan part is equal, whereas we have an isomorphism between the decorations $F(\iota_m)x+F(\iota_{m'})x'\cong \phi_{m,m'}(x,x')$ in an analogous way to \cref{eq:isomorphismdiag} as follows
\begin{displaymath}
\begin{tikzcd}
F(m)\times F(m')\ar[rr,"\phi_{m,m'}"]\ar[d,"F(i_m)\times F(i_{m'})"']\ar[dr,phantom,"\stackrel{\cref{eq:pseudonaturality}}{\cong}"] && F(m+m')\ar[d,equal]\ar[d,bend right=50, phantom,"\cong"']\ar[dl,"F(i_m+i_{m'})"description] \\
F(m+m')\times F(m+m')\ar[r,"\phi_{m+m',m+m'}"']\ar[rr,dashed,bend right=15] & F\left((m+m')+(m+m')\right)\ar[r,"F(\nabla)"'] & F(m+m')
 \end{tikzcd}
\end{displaymath}
We thus constructed isomorphisms $\lE(M)\otimes\lE(M')\cong\lE(M\otimes M')$ for any $M,M'$.
Notice that both sides in the diagram above, combined with the same object $m+m'$ in $\A$, express a coproduct in the Grothendieck category $\inta F$ by formulas \cref{eq:fibrewisecoprod,eq:explicitstructure2}. Therefore the pasted isomorphism is the canonical unique one that commutes with the respective cocones. This is useful when checking associativity for example: there is only one such isomorphism $\lE(M)\otimes\lE(M')\otimes\lE(M'')\simrightarrow\lE(M\ot M'\ot M'')$ for any three structured cospans.

>>>>

% Structured versus decorated cospans
% John Baez, Kenny Courser and Christina Vasilakopoulou
% in LaTeX
% 2020/12/17

@@

The associator for horizontal composition
is formed as follows: for three composable horizontal 1-cells $M_1=(a\to m_1\leftarrow b,x_1\in F(m_1))$, $M_2=(b\to m_2\leftarrow c,x_2\in F(m_2))$ and $M_3=(c\to m_3\leftarrow d,x_3\in F(m_3)),$
the isomorphism $(M_3\odot M_2)\odot M_1\cong M_3\odot (M_2\odot M_1)$ is the isomorphism of cospans 
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0.5) {$a$};
\node (A') at (0,-0.5) {$a$};
\node (B) at (1.5,0.5) {$m_1+_{b}(m_2 +_{c} m_3)$};
\node (C) at (3,0.5) {$d$};
\node (C') at (3,-0.5) {$d$};
\node (D) at (1.5,-0.5) {$(m_1+_{b} m_2)+_{c} m_3$};
\node (E) at (5.5,0.5) {$(x_3 \odot x_2) \odot x_1 \in F(m_1+_{b} (m_2 +_{c} m_3))$};
\node (F) at (5.5,-0.5) {$x_3 \odot (x_2 \odot x_1) \in F((m_1+_{b} m_2)+_{c} m_3)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$$} (B)
(C) edge node[above]{$$} (B)
(A) edge node[left]{$1_a$} (A')
(C) edge node[right]{$1_d$} (C')
(A') edge node {$$} (D)
(C') edge node {$$} (D)
(B) edge node [left] {$\stackrel{\kappa}{\cong}$} (D);
\end{tikzpicture}
\]
together with an isomorphism $F(\kappa)((x_3 \odot x_2) \odot x_1) \cong (x_3 \odot (x_2 \odot x_1))$ between the decorations 
\begin{equation}\label{eq:decoiso}
\scalebox{.7}
{\begin{tikzcd}[ampersand replacement=\&]
\& F(m_1){\times}(F(m_2){\times}F(m_3))\ar[dddrr,phantom,"\stackrel{\cref{eq:omega}}{\cong}"]\ar[r,"1\times\phi_{m_2,m_3}"]\ar[ddd,"\alpha"description] \& F(m_1){\times}F(m_2{+}m_3)\ar[r,"1\times F(\psi)"]\ar[dr,"\phi_{m_1,m_2+m_3}"'] \& F(m_1){\times}F(m_2{+_c}m_3)\ar[d,phantom,"\stackrel{\cref{eq:pseudonaturality}}{\cong}"]\ar[r,"\phi_{m_1,m_2+_cm_3}"] \& F(m_1{+}(m_2{+_c}m_3))\ar[r,"F(\psi)"]\ar[dddr,phantom,"\cong"] \& F(m_1{+_b}(m_2{+_c}m_3))\ar[ddd,"F(\kappa)"] \\
\one\ar[ur,"x_1\times(x_2\times x_3)"]\ar[d,equal] \&\&\& F(m_1{+}(m_2{+}m_3))\ar[ur,"F(1+\psi)"']\ar[d,"F\alpha"] \& \\
\one\ar[dr,"(x_1\times x_2)\times x_3"']\&\&\& F((m_1{+}m_2){+}m_3)\ar[dr,"F(\psi+1)"]\ar[d,phantom,"\stackrel{\cref{eq:pseudonaturality}}{\cong}"] \& \\
\& (F(m_1){\times}F(m_2)){\times}F(m_3)\ar[r,"\phi_{m_1,m_2}\times1"'] \& F(m_1{+}m_2){\times}m_3\ar[ur,"\phi_{m_1+m_2,m_3}"]\ar[r,"F(\psi)\times1"'] \& F(m_1{+_b}m_2){\times} F(m_3)\ar[r,"\phi_{m_1{+_b}m_2,m_3}"'] \& F((m_1{+_b}m_2){+}m_3)\ar[r,"F\psi"'] \& F((m_1{+_b}m_2){+_c}m_3)
\end{tikzcd}}
\end{equation}
where $a$ are the associators of $\Cat$ and $\A$ and $\kappa$ is the canonical isomorphism between two colimits of the same diagram in $\A$. Similarly the right and left unitors can be constructed, and all data satisfy the axioms of a double category---essentially breaking down to those of the cospan double category of $\A$ and axioms of the lax monoidal pseudofunctor $F$. For details of these arguments, see \cite[Theorem~4.1.1]{CourserThesis}.
\end{proof}

@@

\&\& F(m_1+n_1+m_2+n_2)\ar[d,"F(\psi)"] \\

@@

\one \ar[ddrr,dashed,"{(x_2\odot x_1) \otimes (y_2\odot y_1)}"']\ar[r,"x_1{\times}x_2{\times}y_1{\times}y_2"] \& F(m_1){\times}F(m_2){\times}F(n_1){\times}F(n_2)\ar[r,"\phi_{m_1,m_2}\times\phi_{n_1,n_2}"] \& F(m_1{+}m_2){\times}F(n_1{+}n_2) \ar[d, "F(\psi) \times F(\psi)"] \\

@@

\& F(m_1)\times F(n_1)\times F(m_2)\times F(n_2)\ar[dd,"1\times\beta\times1"']\ar[ddrr,phantom,"\stackrel{\cref{eq:thetaa}}{\cong}"]\ar[rr,"\phi_{m_1,n_1}\times\phi_{m_2,n_2}"] \&\& F(m_1+n_1)\times F(m_2+n_2)\ar[rr,"\phi_{m_1+n_1,m_2+n_2}"] \&\& F(m_1+n_1+m_2+n_2)\ar[dd,"F(\psi)"]\ar[ddll,"F(1+\beta+1)"'] \\

@@

where the left-hand side isomorphism combines the pseudoassociativity \cref{eq:omega} and braided monoidal structure \cref{eq:braidedpseudofun} of the pseudofunctor $F$ as seen in detail in the appendix, and the right hand side follows from the universal property of colimits and pseudofunctoriality of $F$.  Similarly, for the units $U_{a+b}$ and $U_a\ot U_b$ it is easy to see that as cospans they are both the identity $1_{a+b} \maps a+b\to a+b\leftarrow a+b:1_{a+b}$, and there is a canonical isomorphism between their trivial decorations \cref{eq:UFCsp}, namely $F(!_{a+b})\phi_0\cong\phi_{a,b}(F(!_a)\times F(!_b))(\phi_0\times\phi_0)$ via the structure isomorphisms for $(\phi,\phi_0)$.

@@

Building towards an isomorphism of the above described double categories, we now define a double functor
\begin{displaymath}
\lE=(\lE_0,\lE_1)\maps _L\lCsp(\inta F) \longrightarrow F\lCsp
\end{displaymath}
which on vertical categories acts as the identity $\lE_0 = \id_{\A}$. Given a horizontal 1-cell in $_L\lCsp(\inta F)$ as in \cref{eq:scsphor1cell}, $\lE_1$ maps it to the decorated cospan $i\maps a\to m\leftarrow b:o$ with decoration $x\in F(m)$,
whereas a 2-morphism of $L$-structured cospans as in \cref{eq:2morphismLCspintaF} is mapped to the underlying map of cospans in $\A$
\begin{displaymath}
\begin{tikzcd}
a\ar[r,"i"]\ar[d,"f"'] & m\ar[d,"k"] & b\ar[l,"o"']\ar[d,"g"] \\
a'\ar[r,"i'"']& m' & b'\ar[l,"o'"]
\end{tikzcd}
\end{displaymath}
in $\A$ along with the map $\tau\maps Fk(x)\to x'$ in $F(m')$ coming from the middle arrow above, namely an $F$-decorated cospan 2-morphism. 
With these mappings, $\lE$ is indeed a double functor as in \cref{def:doublefun}. First of all, both $\lE_0$ and $\lE_1$ are functors that appropriately commute with sources and targets. Moreover, there exist globular 2-isomorphisms $\lE(N) \odot \lE(M) \simrightarrow\lE(N \odot M)$, $U_{\lE m} \simrightarrow\lE(U_m)$
for any pair of composable horizontal 1-cells $M=\cspn{(a,\bot_a)}{(m,x)}{(b,\bot_b)}{}{}$ and $N=\cspn{(b,\bot_b)}{(n,y)}{(c,\bot_c)}{}{}$ and any object $m$ of $_L\lCsp(\inta F)$. Explicitly, $\mathbb{E}(N)\odot \mathbb{E}(M)$ is given as in \cref{eq:dcospanscomposition} via the pushout cospan and decoration
\begin{equation}\label{eq:1}
 \begin{tikzcd}
  & m+_bn & \\
  a\ar[ur] && b,\ar[ul]
 \end{tikzcd}\qquad
  \one\xrightarrow{x\times y}F(m)\times F(n)\xrightarrow{\phi_{m,n}}F(m+n)\xrightarrow{F(\psi)}F(m+_bn)
\end{equation}
whereas first composing $M$ and $N$ in $_L\lCsp(\inta F)$ using pushouts in $\inta F$ as in \cref{eq:globalpushout} produces 
\begin{displaymath}
 \begin{tikzcd}[column sep={.7in,between origins}]
 && (m+_b n,F(v_m)x+_{(\bot_{m+_bn})}F(v_n)y) && \\
 & (m,x) \ar[ur] && (n,y)\ar[ul] \\
 (a,\bot_a)\ar[ur] && (b,\bot_b)\ar[ur]\ar[ul] && (c,\bot_c)\ar[ul]
 \end{tikzcd}
\end{displaymath}
where $v_m,v_n$ are the canonical maps into the pushout.
Under $\lE$ this is mapped to the underlying $\A$-cospan $\cspn{a}{m+_bn}{b}{}{}$ with decoration $F(v_m)x+F(v_n)y$ in the fiber $F(m+_bn)$, since a pushout over an initial object is really a coproduct.  This decoration has a canonical isomorphism to that of \cref{eq:1} due to the following diagram 
\begin{equation}\label{eq:isomorphismdiag}
 \begin{tikzcd}
F(m)\times F(n)\ar[rr,"\phi_{m,n}"]\ar[d,"F(v_m)\times F(v_n)"']\ar[dr,phantom,"\cong"] && F(m+n)\ar[d,"F(\psi)"]\ar[d,bend right=50, phantom,"\cong"']\ar[dl,"F(v_m+v_n)"description] \\
F(m+_bn)\times F(m+_bn)\ar[r,"\phi_{m+_bn,m+_bn}"'] & F\left((m+_bn)+(m+_bn)\right)\ar[r,"F(\nabla)"'] & F(m+_bn)
 \end{tikzcd}
\end{equation}
where the composite of the bottom two arrows is the coproduct $+ \maps F(m+_bn) \times F(m+_bn) \to F(m+_bn)$ as in \cref{eq:explicitstructure1}. The left-hand isomorphism arises from the pseudonaturality of $\phi$ as in \cref{eq:pseudonaturality}, whereas the right-hand isomorphism follows from universal properties and the pseudofunctoriality of $F$.  Regarding identities, the identity horizontal 1-cell $U_m$ in $_L\lCsp(\inta F)$ is $\cspn{(m,\bot_m)}{(m,\bot_m)}{(m,\bot_m)}{}{}$ with $1_m$ as the $A$-component and isomorphisms between initial objects in the fibers. Thus $\lE(U_m)$ is the identity cospan in $A$ together with the `initial decoration' $\bot_m\in F(m)$, and so is $U_{\mathbb{E}(m)}$ as in \cref{eq:UFCsp}.

@@

Since both double categories are symmetric monoidal under the running assumptions (\cref{DC,thm:SC}), we now show that this double isomorphism is a symmetric monoidal one, namely that $\lE=(\lE_0,\lE_1)$ is a symmetric strong monoidal double functor as per \cref{defn:monoidal_double_functor}. Indeed, $\lE_0= \id_\A$ is a symmetric monoidal functor trivially. For $\lE_1$, given two structured cospans $M=(a,\bot_a)\to(m,x)\leftarrow(b,\bot_b)$ and $M'=(a',\bot_{a'})\to(m',x')\leftarrow(b',\bot_{b'})$ in $_L\lCsp(\inta F)$, we compute the appropriate decorated cospans
\begin{align*}
\lE(M)\ot\lE(M')&\;\stackrel{\cref{eq:tensordecoration}}{=}\;a+a'\xrightarrow{i+i'}m+m'\xleftarrow{o+o'}b+b' \textrm{ with }\phi_{m,m'}(x,x')\in F(m+m')
%\one\xrightarrow{x\times x'}F(m)\times F(m')\xrightarrow{\phi_{m,m'}}F(m+m')
\\
\lE(M\ot M')&\;\stackrel{\cref{eq:tensorstuctured}}{=}\;\lE\Big((a+a',\bot_{a+a'})\to\underbrace{(m+m',F(\iota_m)x+F(\iota_{m'})x')}_{\cref{eq:fibrewisecoprod}}\leftarrow(b+b',\bot_{b+b'})\Big) \\
&=a+a'\xrightarrow{i+i'}m+m'\xleftarrow{o+o'}b+b' \textrm{ with }F(\iota_m)x+F(\iota_{m'})x'\in F(m+m')
\end{align*}
The cospan part is equal, whereas we have an isomorphism between the decorations, $F(\iota_m)x+F(\iota_{m'})x'\cong \phi_{m,m'}(x,x')$, given in an analogous way to \cref{eq:isomorphismdiag} as follows:
\begin{displaymath}
\begin{tikzcd}
F(m)\times F(m')\ar[rr,"\phi_{m,m'}"]\ar[d,"F(i_m)\times F(i_{m'})"']\ar[dr,phantom,"\stackrel{\cref{eq:pseudonaturality}}{\cong}"] && F(m+m')\ar[d,equal]\ar[d,bend right=50, phantom,"\cong"']\ar[dl,"F(i_m+i_{m'})"description] \\
F(m+m')\times F(m+m')\ar[r,"\phi_{m+m',m+m'}"'] & F\left((m+m')+(m+m')\right)\ar[r,"F(\nabla)"'] & F(m+m')
 \end{tikzcd}
\end{displaymath}
where the composite of the bottom two arrows is the coproduct operation in $F(m+m')$.
We thus have constructed an isomorphism $\lE(M)\otimes\lE(M') \simrightarrow \lE(M\otimes M')$ for each $M,M'$. Notice that both sides in the diagram above, combined with the same object $m+m'$ in $\A$, express a coproduct in the Grothendieck category $\inta F$ by formulas \cref{eq:fibrewisecoprod,eq:explicitstructure2}. Therefore the isomorphism $F(\iota_m)x+F(\iota_{m'})x '\cong \phi_{m,m'}(x,x')$ given by this diagram is the unique one that commutes with the respective cocones.  This is useful when checking associativity for example: there is only one such isomorphism $\lE(M)\otimes\lE(M')\otimes\lE(M'')\simrightarrow\lE(M\ot M'\ot M'')$ for any three structured cospans.
